---
title: "Prep flood ensemble model"
output: html_document
date: '2022-07-11'
editor_options: 
  chunk_output_type: console
---

These are 5 year return periods. Note: the website with these data added an extra zero after the rp (which is confusing).

https://files.wri.org/d8/s3fs-public/aqueduct-floods-methodology.pdf
Winsemius et al. (2016) argue in their supplementary materials that a bias correction on the climate scenarios is required because the input climate datasets
are only bias-corrected based on daily statistics. They thus may still contain
bias in the autocorrelation of rainfall in time, at both short and long (yearly) time
scales, as demonstrated by Johnson et al. (2011). Therefore, an additional bias
correction was required. In Winsemius et al. (2016), this bias correction was applied by using a model-model diference in flood risk rather than by applying a
bias correction within the hazard simulations themselves before performing the
hazard map downscaling procedure. For Aqueduct Floods, we performed an additive correction method within the space of the extreme value distributions of
the flood volumes. The extreme volumes obtained from the run forced with the
observation-based EUWATCH dataset (1960–99) were assumed as the presentday values, and their changes in future flood volumes were calculated, for all
GCMs, from diferences between the values in future (RCP and GCM combination) and historical (GCM) conditions. The procedure to obtain corrected future
volumes, “future corrected,” is summarized in the following equation:

future_corrected = present_watch + (future_gcm - present_gcm)

with the present_watch representing present-day values based on the
EUWATCH run, and present_gcm and future_gcm indicating values obtained
from GCM runs in historical (under present-day greenhouse gas concentration
forcing) and future (for various RCP scenarios) simulation periods. 

Aqueduct Floods now includes two hazard components,
riverine and coastal foods. Both hazards are represented
by global scale layers of inundation depth at 30″ × 30″
resolution, with diferent layers representing inundation
depths for diferent annual average probabilities of occurrence. The hazard layers have been simulated without
considering the presence of food protection. Note that
this does not mean that food protection is not included in
our risk computation. Instead of including it in the computations of riverine and coastal food hazard, we include
the efects of food protection in the risk calculations by
assuming zero damage below the assumed standard of
protection (see Section 6).

Based on PCR-GLOBWB output,
we then applied extreme value statistics to derive the
foodplain water volumes per grid cell for several fooding
return periods (2, 5, 10, 25, 50, 100, 250, 500, and 1,000 
4 |
years) for the current time (based on 1960–99 simulation)
and future climate (2010–49, 2030–69, and 2060–99).
These are then used as input to a volume spreading food
model (Winsemius et al. 2013) to convert the 5′ × 5′ food
volumes into maps of high-resolution inundation depth at
a resolution of 30″ × 30″. We provide additional documentation on PCR-GLOBWB (version 2) and GLOFRIS
model details, input data, and bias correction processes in
Appendix A.1.1.
for future conditions we used the Inter-sectoral Impact Model Intercomparison
Project (ISI-MIP) forcing data (Hempel et al. 2013) for the periods 2030 (by running over the period 2010–49) and 2050 (by running over the period 2030–69),
as well as 2080 (2060–99). Note that the exposure data for current conditions
represent the situation in 2010 (not 1960–99). 

Data from here: http://wri-projects.s3.amazonaws.com/AqueductFloodTool/download/v2/index.html
```{r} 
library(sp)
library(raster)

yr_list <- c("2030_rp00005.tif", "2050_rp00005.tif", "2080_rp00005.tif")

```

Do the rcp4p5 ones first:
```{r}
for (yr in yr_list){ #yr=yr_list[1]

  rasts <- list.files("/home/shares/aquaculture/aquaculture_climatechange_usa/raw_data/flood/river_flood", pattern=yr, full=TRUE)
   rasts <- grep("rcp4p5", rasts, value=TRUE)
   
  rast_stack <- raster::stack(rasts)
  #plot(rast_stack[[1]])
  num_rasts <- length(rasts)
  cat(sprintf("analyzing %s, n = %s", yr, num_rasts))
  calc(rast_stack, fun=mean, filename=sprintf("/home/shares/aquaculture/aquaculture_climatechange_usa/prep_data/freshwater/flood/ensembles_river/inunriver_rcp4p5_ensemble_%s", yr), overwrite=TRUE) }


tmp <- raster("/home/shares/aquaculture/aquaculture_climatechange_usa/prep_data/freshwater/flood/ensembles_river/inunriver_rcp4p5_ensemble_2050_rp00005.tif")  
 plot(area(tmp))
```



Do the rcp8p5 ones next:
```{r}
for (yr in yr_list){ #yr=yr_list[1]

  rasts <- list.files("/home/shares/aquaculture/aquaculture_climatechange_usa/raw_data/flood/river_flood", pattern=yr, full=TRUE)
  rasts <- grep("rcp8p5", rasts, value=TRUE)
  rast_stack <- raster::stack(rasts)
  #plot(rast_stack[[1]])
  num_rasts <- length(rasts)
  cat(sprintf("analyzing %s, n = %s", yr, num_rasts))
  calc(rast_stack, fun=mean, filename=sprintf("/home/shares/aquaculture/aquaculture_climatechange_usa/prep_data/freshwater/flood/ensembles_river/inunriver_rcp8p5_ensemble_%s", yr), overwrite=TRUE) }


tmp <- plot(raster("/home/shares/aquaculture/aquaculture_climatechange_usa/prep_data/freshwater/flood/ensembles_river/inunriver_rcp8p5_2050_rp00005.tif"))  
 
```