---
title: "Figure: change in average pressures barplot"
output: html_document
date: "2023-09-11"
editor_options: 
  chunk_output_type: console
---
Be sure to run step0 before doing this!

Determine mean pressure (for each individual pressure) for each taxa in its high probability location:

```{r setup, include=FALSE}
library(here)
library(terra)
library(tidyverse)
library(sf)
library(raster)

#raster template
template <- raster()
template <- raster(xmn=-6500000, xmx=2500000, ymn=100000, ymx=6200000)

crs(template) <- CRS('+init=EPSG:5070')
res(template) <- 5000

#writeRaster(template, here("spatial/template.tif"))
```

```{r}
list.files("/home/shares/aquaculture/aquaculture_climatechange_usa/prep_data/freshwater/Tavg_taxa_prs/rescaled_yearly_crs")

combos <- read_csv(here("pressure_calcs/filepaths_for_pressure_combos.csv"), na="NA")

#list of taxa we will evaluate
taxa <- c("trout", "tilapia", "carp", "catfish", "sturgeon", "bass", "perch")

```
list of scenario years we want to run (pressures for each category will be combined):
```{r}

yr_scens <- unique(combos$year_scenario)
yr_scens <- yr_scens[!is.na(yr_scens )]

table(combos$year_scenario) # check that all have 4 layers!

```

```{r}

for(yr_scen in yr_scens){
#yr_scen = yr_scens[5]

for(taxon in taxa){ # taxon="trout"
  
taxa_location <- rast(sprintf("/home/shares/aquaculture/aquaculture_climatechange_usa/prep_data/freshwater/mapping_aquaculture/estimated_farm_locations/%s.tif", taxon))  
  
rast_paths <- filter(combos, year_scenario %in% yr_scen) %>%
  mutate(file_name = ifelse(pressure == "tmp_taxa", paste0(taxon, file_name), file_name)) %>%
  mutate(path = sprintf("%s/%s.tif", file_path, file_name)) %>%
  pull(path)

rast_stack <- rast(rast_paths)

taxa_summary <- terra::zonal(rast_stack, taxa_location, fun="mean", na.rm=TRUE) %>%
rename_with(~ gsub(taxon, "species", .)) %>%
  pivot_longer(cols = -species, 
               names_to = "variable", 
               values_to = "value") %>%
  rename(presence = species)

taxa_summary$yr_scen <- yr_scen
taxa_summary$taxa <- taxon

write_csv(taxa_summary, sprintf("figs/pressure_by_taxa_scenario_datasets/%s_%s.csv", taxon, yr_scen))

cat("finished ", taxon, " : ", yr_scen, ", with N rasters = ", dim(rast_stack)[3], "\n")

}
}
```


Combine each taxa file into a single dataset and clean the data.
```{r}

files_list <- list.files("figs/pressure_by_taxa_scenario_datasets/", full=TRUE)

combined_data <- files_list %>%
  set_names() %>%
  map_df(~read_csv(.))

combined_data2 <- combined_data %>%
  mutate(variable = str_replace(variable, "^wb_.*", "wet bulb"),
         variable = str_replace(variable, "^inuncoast_.*", "flood"),
         variable = str_replace(variable, "^CMD_.*", "water limitation"),
         variable = str_replace(variable, "^species_.*", "temperature")) %>%
     #    taxa = str_replace(taxa, "bass", "large mouth bass"),
      #   taxa = str_replace(taxa, "trout", "trout spp.")) %>%
  filter(presence == 1) %>%
  dplyr::select(-presence)  %>%
  separate(yr_scen, into=c("time", "scenario"), sep = "_") %>%
  mutate(variable = factor(variable, levels = c("temperature", "wet bulb", "flood", "water limitation")))



current <- combined_data2 %>%
  filter(time == "current") %>%
  rename(current_value = value) %>%
  dplyr::select(-time)

future_change <- combined_data2 %>%
  filter(time != "current") %>%
  rename(proj_value = value) %>%
  left_join(current, by=c("scenario", "taxa", "variable")) %>%
  mutate(proj_change = proj_value - current_value)

```

Plotting the data
```{r}
#install.packages("envalysis")
library(envalysis)
#install.packages("patchwork")
library(patchwork)

ssp2 <- ggplot(data = filter(future_change, time=="near-future" & scenario == "ssp245"), aes(x = taxa, y = proj_change, fill = variable)) +
  geom_bar(stat = "identity", position = "stack") +
  coord_flip() +
  theme_publish() +
  labs(title = "SSP 2",
       y = "Projected change in pressure",
       x = "") +
  scale_x_discrete(limits = rev(levels(future_change$taxa))) + 
  scale_fill_manual(values = c("wet bulb" = "#C44221", 
                               "temperature" = "#ED6F2E", 
                               "flood" = "#22AECC", 
                               "water limitation" = "#D2AE7B")) +
  theme(
    legend.position = c(0.9, -0.06), 
    legend.title = element_blank(),
    legend.justification = c(1, 0),
    legend.box.just = "right",
    axis.title.x = element_text(color = "transparent"),  # Make x-axis title transparent
    axis.text.x = element_text(color = "transparent"),   # Make x-axis text transparent
    axis.ticks.x = element_line(color = "transparent"),  # Make x-axis ticks transparent
    axis.line.x = element_line(color = "transparent")    # Make x-axis line transparent
  ) +
  guides(fill = guide_legend(nrow = 4, keywidth = 1, keyheight = 1)) +
  geom_hline(yintercept = 0, linetype = "solid", color = "gray", linewidth=2, alpha = 0.5) + 
  geom_hline(yintercept = 0, linetype = "solid", color = "black", linewidth=0.5, alpha = 0.5) +
  ylim(-0.1, 0.55) + 
  theme(plot.margin = margin(b = -2, unit = "cm"))


#ggsave(here("figs/output/projected_change_ssp2.png"), width = 5, height=3, units="in")

ssp5 <- ggplot(data = filter(future_change, time=="near-future" & scenario == "ssp585"), aes(x = taxa, y = proj_change, fill = variable)) +
  geom_bar(stat = "identity", position = "stack") +
  coord_flip() +
  theme_publish() +
  labs(title = "SSP 5",
       y = "Projected change in pressure",
       x = "") +
  scale_x_discrete(limits = rev(levels(future_change$taxa))) + # To have 0 in the middle
  scale_fill_manual(values = c("wet bulb" = "#C44221", 
                               "temperature" = "#ED6F2E", 
                               "flood" = "#22AECC", 
                               "water limitation" = "#D2AE7B")) +
  theme(legend.position = "none") +  # Remove legend
  geom_hline(yintercept = 0, linetype = "solid", color = "gray", linewidth=1.5, alpha = 0.5) + 
  geom_hline(yintercept = 0, linetype = "solid", color = "black", linewidth=0.5, alpha = 0.5) + 
  ylim(-0.1, 0.55)+ 
  theme(plot.margin = margin(t = -2, unit = "cm"))


#ggsave(here("figs/output/projected_change_ssp5.png"), width = 5, height=3, units="in")

stacked <- ssp2/ssp5
ggsave("figs/output/fig6.png", width = 5, height = 6)

```

Ended here, so following might need to be revised.

Get cumulative pressures in the same way (cumulative pressures are multiplicative so viewing individual pressures isn't exactly intuitive).

```{r}

for(yr_scen in yr_scens){
#yr_scen = yr_scens[4]
  
  for(taxon in taxa){ 
  # taxon = "trout"
rast_paths <- filter(combos, year_scenario %in% yr_scen) %>%
  mutate(file_name = ifelse(pressure == "tmp_taxa", paste0(taxon, file_name), file_name)) %>%
  mutate(path = sprintf("%s/%s.tif", file_path, file_name)) %>%
  pull(path)
cat(length(rast_paths))

rast_stack <- rast(rast_paths)
rast_stack[is.na(rast_stack)] <- 0
rast_stack_inverse <- 1-rast_stack
rast_prod <- app(rast_stack_inverse, fun=prod)
#rast_gm_inverse <- rast_prod^(1/dim(rast_stack)[3])
rast_gm <- 1-rast_prod
#writeRaster(rast_gm, sprintf("/home/shares/aquaculture/aquaculture_climatechange_usa/prep_data/freshwater/combined_pressures/taxa_pressures/%s_%s.tif", yr_scen, taxon), overwrite=TRUE)
}}
```


