---
title: "Taxa biological pressures"
output: html_document
date: "2023-10-03"
editor_options: 
  chunk_output_type: console
---
This creates maps for each taxa for current and future projected biological pressures based on thermal tolerance of each taxa.
```{r}
library(terra)
library(tidyverse)
library(RColorBrewer)

```

```{r}
US <- terra::rast(here("spatial/USA_raster.tif"))

pressure_prep <- function(path = "/home/shares/aquaculture/aquaculture_climatechange_usa/prep_data/freshwater/combined_pressures/current_ssp245_no_taxa_specific.tif"){

tmp_raster <- terra::rast(path)
tmp_raster[is.na(tmp_raster)] <- 0
tmp_raster <- tmp_raster*US
tmp_raster
}


locs <- "/home/shares/aquaculture/aquaculture_climatechange_usa/prep_data/freshwater/Tavg_taxa_prs/rescaled_yearly_crs/"

#bass
current_bass_ssp245 <- pressure_prep(path =  file.path(locs, "bass_Tavg_ssp245_2001-2020.tif"))
current_bass_ssp585 <- pressure_prep(path =  file.path(locs, "bass_Tavg_ssp585_2001-2020.tif"))
near_bass_ssp245 <- pressure_prep(path =  file.path(locs, "bass_Tavg_ssp245_2041-2060.tif"))
near_bass_ssp585 <- pressure_prep(path =  file.path(locs, "bass_Tavg_ssp585_2041-2060.tif"))

#carp
current_carp_ssp245 <- pressure_prep(path =  file.path(locs, "carp_Tavg_ssp245_2001-2020.tif"))
current_carp_ssp585 <- pressure_prep(path =  file.path(locs, "carp_Tavg_ssp585_2001-2020.tif"))
near_carp_ssp245 <- pressure_prep(path =  file.path(locs, "carp_Tavg_ssp245_2041-2060.tif"))
near_carp_ssp585 <- pressure_prep(path =  file.path(locs, "carp_Tavg_ssp585_2041-2060.tif"))

#catfish
current_catfish_ssp245 <- pressure_prep(path =  file.path(locs, "catfish_Tavg_ssp245_2001-2020.tif"))
current_catfish_ssp585 <- pressure_prep(path =  file.path(locs, "catfish_Tavg_ssp585_2001-2020.tif"))
near_catfish_ssp245 <- pressure_prep(path =  file.path(locs, "catfish_Tavg_ssp245_2041-2060.tif"))
near_catfish_ssp585 <- pressure_prep(path =  file.path(locs, "catfish_Tavg_ssp585_2041-2060.tif"))

#perch
current_perch_ssp245 <- pressure_prep(path =  file.path(locs, "perch_Tavg_ssp245_2001-2020.tif"))
current_perch_ssp585 <- pressure_prep(path =  file.path(locs, "perch_Tavg_ssp585_2001-2020.tif"))
near_perch_ssp245 <- pressure_prep(path =  file.path(locs, "perch_Tavg_ssp245_2041-2060.tif"))
near_perch_ssp585 <- pressure_prep(path =  file.path(locs, "perch_Tavg_ssp585_2041-2060.tif"))


#sturgeon
current_sturgeon_ssp245 <- pressure_prep(path =  file.path(locs, "sturgeon_Tavg_ssp245_2001-2020.tif"))
current_sturgeon_ssp585 <- pressure_prep(path =  file.path(locs, "sturgeon_Tavg_ssp585_2001-2020.tif"))
near_sturgeon_ssp245 <- pressure_prep(path =  file.path(locs, "sturgeon_Tavg_ssp245_2041-2060.tif"))
near_sturgeon_ssp585 <- pressure_prep(path =  file.path(locs, "sturgeon_Tavg_ssp585_2041-2060.tif"))
                      
#tilapia
current_tilapia_ssp245 <- pressure_prep(path =  file.path(locs, "tilapia_Tavg_ssp245_2001-2020.tif"))
current_tilapia_ssp585 <- pressure_prep(path =  file.path(locs, "tilapia_Tavg_ssp585_2001-2020.tif"))
near_tilapia_ssp245 <- pressure_prep(path =  file.path(locs, "tilapia_Tavg_ssp245_2041-2060.tif"))
near_tilapia_ssp585 <- pressure_prep(path =  file.path(locs, "tilapia_Tavg_ssp585_2041-2060.tif"))

#trout
current_trout_ssp245 <- pressure_prep(path =  file.path(locs, "trout_Tavg_ssp245_2001-2020.tif"))
current_trout_ssp585 <- pressure_prep(path =  file.path(locs, "trout_Tavg_ssp585_2001-2020.tif"))
near_trout_ssp245 <- pressure_prep(path =  file.path(locs, "trout_Tavg_ssp245_2041-2060.tif"))
near_trout_ssp585 <- pressure_prep(path =  file.path(locs, "trout_Tavg_ssp585_2041-2060.tif"))

```

```{r calculate differences}


# the crit point is the maximum plotting value
crit_min_plot <- -0.5
crit_max_plot <- 0.5

diff_function <- function(current_raster = cumulative_current_ssp245, future_raster = cumulative_near_ssp245,
                          crit_min = crit_min_plot, crit_max = crit_max_plot){
  
  projected_change <- future_raster - current_raster
  projected_change <- terra::ifel(projected_change > crit_max, crit_max, projected_change)
  projected_change <- terra::ifel(projected_change < crit_min, crit_min, projected_change)
}

bass_dif_near_ssp245 <- diff_function(future_raster = near_bass_ssp245, current_raster = current_bass_ssp245)
bass_dif_near_ssp585 <- diff_function(future_raster = near_bass_ssp585, current_raster = current_bass_ssp585)

catfish_dif_near_ssp245 <- diff_function(future_raster = near_catfish_ssp245, current_raster = current_catfish_ssp245)
catfish_dif_near_ssp585 <- diff_function(future_raster = near_catfish_ssp585, current_raster = current_catfish_ssp585)

carp_dif_near_ssp245 <- diff_function(future_raster = near_carp_ssp245, current_raster = current_carp_ssp245)
carp_dif_near_ssp585 <- diff_function(future_raster = near_carp_ssp585, current_raster = current_carp_ssp585)

tilapia_dif_near_ssp245 <- diff_function(future_raster = near_tilapia_ssp245, current_raster = current_tilapia_ssp245)
tilapia_dif_near_ssp585 <- diff_function(future_raster = near_tilapia_ssp585, current_raster = current_tilapia_ssp585)

perch_dif_near_ssp245 <- diff_function(future_raster = near_perch_ssp245, current_raster = current_perch_ssp245)
perch_dif_near_ssp585 <- diff_function(future_raster = near_perch_ssp585, current_raster = current_perch_ssp585)

trout_dif_near_ssp245 <- diff_function(future_raster = near_trout_ssp245, current_raster = current_trout_ssp245)
trout_dif_near_ssp585 <- diff_function(future_raster = near_trout_ssp585, current_raster = current_trout_ssp585)

sturgeon_dif_near_ssp245 <- diff_function(future_raster = near_sturgeon_ssp245, current_raster = current_sturgeon_ssp245)
sturgeon_dif_near_ssp585 <- diff_function(future_raster = near_sturgeon_ssp585, current_raster = current_sturgeon_ssp585)

```

```{r convert to points}
#convert to points

dataframe_convert <- function(r=wet_bulb_current_ssp245){
  temp <- as.data.frame(r, xy=TRUE)
  colnames(temp) <- c("x", "y", "value")
  temp
}

pt_bass_current <- dataframe_convert(r=current_bass_ssp245)
pt_bass_near_ssp245 <- dataframe_convert(bass_dif_near_ssp245)
pt_bass_near_ssp585 <- dataframe_convert(bass_dif_near_ssp585)

pt_catfish_current <- dataframe_convert(r=current_catfish_ssp245)
pt_catfish_near_ssp245 <- dataframe_convert(catfish_dif_near_ssp245)
pt_catfish_near_ssp585 <- dataframe_convert(catfish_dif_near_ssp585)

pt_carp_current <- dataframe_convert(r=current_carp_ssp245)
pt_carp_near_ssp245 <- dataframe_convert(carp_dif_near_ssp245)
pt_carp_near_ssp585 <- dataframe_convert(carp_dif_near_ssp585)

pt_tilapia_current <- dataframe_convert(r=current_tilapia_ssp245)
pt_tilapia_near_ssp245 <- dataframe_convert(tilapia_dif_near_ssp245)
pt_tilapia_near_ssp585 <- dataframe_convert(tilapia_dif_near_ssp585)

pt_perch_current <- dataframe_convert(r=current_perch_ssp245)
pt_perch_near_ssp245 <- dataframe_convert(perch_dif_near_ssp245)
pt_perch_near_ssp585 <- dataframe_convert(perch_dif_near_ssp585)

pt_trout_current <- dataframe_convert(r=current_trout_ssp245)
pt_trout_near_ssp245 <- dataframe_convert(trout_dif_near_ssp245)
pt_trout_near_ssp585 <- dataframe_convert(trout_dif_near_ssp585)

pt_sturgeon_current <- dataframe_convert(r=current_sturgeon_ssp245)
pt_sturgeon_near_ssp245 <- dataframe_convert(sturgeon_dif_near_ssp245)
pt_sturgeon_near_ssp585 <- dataframe_convert(sturgeon_dif_near_ssp585)

```

# Get together plotting functions and such
```{r make some maps- drought}

## Get current color scheme
library(RColorBrewer)
current_cols <- c("#f0eada", brewer.pal("RdPu", n=9)[2:9])

library(wesanderson)
pal <- wes_palette("Zissou1")
change_cols <- c("#3A9AB2", "#79B6C5", "#ABDDDE", "#f0eada", "#E1AF00", "#FA8400", "#972D14") 
                 #"#F22300")

#change_cols <- rev(brewer.pal("Spectral", n=11))

## Get boundaries
#install.packages("tigris")
library(rnaturalearth)
library(tigris)
us_county <- st_read(here("spatial/state_county_shp/state_county_shp.shp"))

us_outline <- ne_countries(country = "United States of America", returnclass = "sf", scale="large")
us_outline <- st_transform(us_outline, terra::crs(trout_dif_near_ssp245))
plot(us_outline$geometry)

us_states <- ne_states(country = "United States of America", returnclass = "sf")
us_states <- st_transform(us_states, terra::crs(trout_dif_near_ssp245))

alaska <- us_states[us_states$name == "Alaska", ]
plot(alaska$geometry)

hawaii <- us_states[us_states$name == "Hawaii", ]
plot(hawaii$geometry)

## create functions for different plots


## Combining mainland, Alaska and Hawaii
M_A_H <- function(plot1, plot2, plot3) {
combined <- ggdraw() +
  draw_plot(plot1) +
  draw_plot(plot2, width=0.35, height=0.40,  x = 0, y = 0)+
  draw_plot(plot3,  width=0.2, height=0.3,  x = .3, y = .04)+
  theme(panel.background = element_rect(fill = '#b9c9c9', color=NA, size = 0.5))
combined
}

## Current pressures plot that can be generated for each taxa!
current_plot <- function(taxa = "bass", habitat_only = FALSE){ #habitat_only = TRUE
  
  pt_current <- paste0("pt_", taxa, "_current")
  
   habitat <- terra::rast(sprintf("/home/shares/aquaculture/aquaculture_climatechange_usa/prep_data/freshwater/mapping_aquaculture/estimated_farm_locations/%s.tif", taxa))
  habitat <- terra::ifel(habitat==0, 1, NA)
  habitat_df <- dataframe_convert(habitat)
 
  
mainland <- ggplot() +
 geom_tile(data = get(pt_current), aes(x = x, y = y, fill = value)) +
      theme_classic()+ #ggtitle("   Current drought") + 
  scale_fill_gradientn(colours = current_cols, limit=c(0, 1)) +
  xlim(-3250000, 2250000) + ylim(125000, 3300000) +
  theme_void() + 
  theme(legend.position = "none") +
  geom_sf(data=us_outline, fill=NA, color = "gray50", size=0.1, alpha=0.5)   


 alaska <- ggplot() +
 geom_tile(data = get(pt_current) , aes(x = x, y = y, fill = value)) +
  theme_classic() +
  scale_fill_gradientn(colours = current_cols, limit=c(0, 1)) +
  xlim(-5000000, -2150000)+ylim(3900000, 6200000) +
  theme_void() +
  theme(legend.position = "none") +
  geom_sf(data=us_outline, fill=NA, color = "gray50", size=0.1, alpha=0.5) 


  hawaii <- ggplot() +
 geom_tile(data = get(pt_current), aes(x = x, y = y, fill = value)) +
  theme_classic() + 
  scale_fill_gradientn(colours = current_cols, limit=c(0, 1)) +
  xlim(-6300000, -5950000) +
  ylim(1500000, 2100000) +
  theme_void() +
  theme(legend.position = "none") + 
  theme(legend.position = "none") +
  geom_sf(data=us_outline, fill=NA, color = "gray50", size=0.1, alpha = 0.5) 
#  theme(panel.border = element_rect(colour = "gray", fill=NA, size=1))

 if(habitat_only){
  mainland <- mainland +
    geom_tile(data = habitat_df, aes(x=x, y=y), fill="gray")+
  geom_sf(data=us_outline, fill=NA, color = "gray50", size=0.1, alpha=0.5) 

     hawaii <- hawaii +
    geom_tile(data = habitat_df, aes(x=x, y=y), fill="gray")+
  geom_sf(data=us_outline, fill=NA, color = "gray50", size=0.1, alpha=0.5) 
     
      alaska <- alaska +
    geom_tile(data = habitat_df, aes(x=x, y=y), fill="gray")+
  geom_sf(data=us_outline, fill=NA, color = "gray50", size=0.1, alpha=0.5) 

}
  
  
return(list(mainland = mainland, alaska = alaska, hawaii = hawaii))
  
}


## Change in pressures plot that can be generated for each taxa!
change_plot <- function(taxa = "bass", ssp="ssp245", habitat_only = FALSE){
  
  pt_change <- paste0("pt_", taxa, "_near_", ssp)

   habitat <- terra::rast(sprintf("/home/shares/aquaculture/aquaculture_climatechange_usa/prep_data/freshwater/mapping_aquaculture/estimated_farm_locations/%s.tif", taxa))
  habitat <- terra::ifel(habitat==0, 1, NA)
  habitat_df <- dataframe_convert(habitat)
  
  
mainland <- ggplot() +
 geom_tile(data = get(pt_change) , aes(x = x, y = y, fill = value)) +
      theme_classic()+ #ggtitle("   Current drought") + 
  scale_fill_gradientn(colours = change_cols, limit=c(crit_min_plot, crit_max_plot)) +
  xlim(-3250000, 2250000) + ylim(125000, 3300000) +
  theme_void() + 
  theme(legend.position = "none") +
  geom_sf(data=us_outline, fill=NA, color = "gray50", size=0.1, alpha=0.5)   


 alaska <- ggplot() +
 geom_tile(data = get(pt_change), aes(x = x, y = y, fill = value)) +
  theme_classic() +
 scale_fill_gradientn(colours = change_cols, limit=c(crit_min_plot, crit_max_plot)) +
  xlim(-5000000, -2150000)+ylim(3900000, 6200000) +
  theme_void() +
  theme(legend.position = "none") +
  geom_sf(data=us_outline, fill=NA, color = "gray50", size=0.1, alpha=0.5) 
#  theme(panel.border = element_rect(colour = "gray", fill=NA, size=1)) 
 

   hawaii <- ggplot() +
 geom_tile(data = get(pt_change), aes(x = x, y = y, fill = value)) +
  theme_classic() + 
 scale_fill_gradientn(colours = change_cols, limit=c(crit_min_plot, crit_max_plot)) +
  xlim(-6300000, -5950000) +
  ylim(1500000, 2100000) +
  theme_void() +
  theme(legend.position = "none") + 
  theme(legend.position = "none") +
  geom_sf(data=us_outline, fill=NA, color = "gray50", size=0.1, alpha = 0.5) 
#  theme(panel.border = element_rect(colour = "gray", fill=NA, size=1))
 
  if(habitat_only){
  mainland <- mainland +
    geom_tile(data = habitat_df, aes(x=x, y=y), fill="gray")+
  geom_sf(data=us_outline, fill=NA, color = "gray50", size=0.1, alpha=0.5) 

     hawaii <- hawaii +
    geom_tile(data = habitat_df, aes(x=x, y=y), fill="gray")+
  geom_sf(data=us_outline, fill=NA, color = "gray50", size=0.1, alpha=0.5) 
     
      alaska <- alaska +
    geom_tile(data = habitat_df, aes(x=x, y=y), fill="gray")+
  geom_sf(data=us_outline, fill=NA, color = "gray50", size=0.1, alpha=0.5) 

}
  
    
# combine mainland, alaska, hawaii
return(list(mainland = mainland, alaska = alaska, hawaii = hawaii))

}   


```



First do all US
```{r}

# bass
bass_current_plots <- current_plot(taxa = "bass")
bass_current <- M_A_H(bass_current_plots$mainland, bass_current_plots$alaska, bass_current_plots$hawaii)

bass_near_ssp245_plots <- change_plot(taxa = "bass", ssp="ssp245")
bass_near_ssp245 <- M_A_H(bass_near_ssp245_plots$mainland, bass_near_ssp245_plots$alaska, bass_near_ssp245_plots$hawaii)

bass_near_ssp585_plots <- change_plot(taxa = "bass", ssp="ssp585")
bass_near_ssp585 <- M_A_H(bass_near_ssp585_plots$mainland, bass_near_ssp585_plots$alaska, bass_near_ssp585_plots$hawaii)
 

# catfish
catfish_current_plots <- current_plot(taxa = "catfish")
catfish_current <- M_A_H(catfish_current_plots$mainland, catfish_current_plots$alaska, catfish_current_plots$hawaii)

catfish_near_ssp245_plots <- change_plot(taxa = "catfish", ssp="ssp245")
catfish_near_ssp245 <- M_A_H(catfish_near_ssp245_plots$mainland, catfish_near_ssp245_plots$alaska, catfish_near_ssp245_plots$hawaii)

catfish_near_ssp585_plots <- change_plot(taxa = "catfish", ssp="ssp585")
catfish_near_ssp585 <- M_A_H(catfish_near_ssp585_plots$mainland, catfish_near_ssp585_plots$alaska, catfish_near_ssp585_plots$hawaii)
 
# carp
carp_current_plots <- current_plot(taxa = "carp")
carp_current <- M_A_H(carp_current_plots$mainland, carp_current_plots$alaska, carp_current_plots$hawaii)

carp_near_ssp245_plots <- change_plot(taxa = "carp", ssp="ssp245")
carp_near_ssp245 <- M_A_H(carp_near_ssp245_plots$mainland, carp_near_ssp245_plots$alaska, carp_near_ssp245_plots$hawaii)

carp_near_ssp585_plots <- change_plot(taxa = "carp", ssp="ssp585")
carp_near_ssp585 <- M_A_H(carp_near_ssp585_plots$mainland, carp_near_ssp585_plots$alaska, carp_near_ssp585_plots$hawaii)
 
# tilapia
tilapia_current_plots <- current_plot(taxa = "tilapia")
tilapia_current <- M_A_H(tilapia_current_plots$mainland, tilapia_current_plots$alaska, tilapia_current_plots$hawaii)

tilapia_near_ssp245_plots <- change_plot(taxa = "tilapia", ssp="ssp245")
tilapia_near_ssp245 <- M_A_H(tilapia_near_ssp245_plots$mainland, tilapia_near_ssp245_plots$alaska, tilapia_near_ssp245_plots$hawaii)

tilapia_near_ssp585_plots <- change_plot(taxa = "tilapia", ssp="ssp585")
tilapia_near_ssp585 <- M_A_H(tilapia_near_ssp585_plots$mainland, tilapia_near_ssp585_plots$alaska, tilapia_near_ssp585_plots$hawaii)
 

# perch
perch_current_plots <- current_plot(taxa = "perch")
perch_current <- M_A_H(perch_current_plots$mainland, perch_current_plots$alaska, perch_current_plots$hawaii)

perch_near_ssp245_plots <- change_plot(taxa = "perch", ssp="ssp245")
perch_near_ssp245 <- M_A_H(perch_near_ssp245_plots$mainland, perch_near_ssp245_plots$alaska, perch_near_ssp245_plots$hawaii)

perch_near_ssp585_plots <- change_plot(taxa = "perch", ssp="ssp585")
perch_near_ssp585 <- M_A_H(perch_near_ssp585_plots$mainland, perch_near_ssp585_plots$alaska, perch_near_ssp585_plots$hawaii)
 

# trout
trout_current_plots <- current_plot(taxa = "trout")
trout_current <- M_A_H(trout_current_plots$mainland, trout_current_plots$alaska, trout_current_plots$hawaii)

trout_near_ssp245_plots <- change_plot(taxa = "trout", ssp="ssp245")
trout_near_ssp245 <- M_A_H(trout_near_ssp245_plots$mainland, trout_near_ssp245_plots$alaska, trout_near_ssp245_plots$hawaii)

trout_near_ssp585_plots <- change_plot(taxa = "trout", ssp="ssp585")
trout_near_ssp585 <- M_A_H(trout_near_ssp585_plots$mainland, trout_near_ssp585_plots$alaska, trout_near_ssp585_plots$hawaii)

# sturgeon
sturgeon_current_plots <- current_plot(taxa = "sturgeon")
sturgeon_current <- M_A_H(sturgeon_current_plots$mainland, sturgeon_current_plots$alaska, sturgeon_current_plots$hawaii)

sturgeon_near_ssp245_plots <- change_plot(taxa = "sturgeon", ssp="ssp245")
sturgeon_near_ssp245 <- M_A_H(sturgeon_near_ssp245_plots$mainland, sturgeon_near_ssp245_plots$alaska, sturgeon_near_ssp245_plots$hawaii)

sturgeon_near_ssp585_plots <- change_plot(taxa = "sturgeon", ssp="ssp585")
sturgeon_near_ssp585 <- M_A_H(sturgeon_near_ssp585_plots$mainland, sturgeon_near_ssp585_plots$alaska, sturgeon_near_ssp585_plots$hawaii)

```


Next do only their suspected ranges:
```{r}

# bass
bass_current_plots_v2 <- current_plot(taxa = "bass", habitat_only = TRUE)
bass_current_v2 <- M_A_H(bass_current_plots_v2$mainland, bass_current_plots_v2$alaska, bass_current_plots_v2$hawaii)

bass_near_ssp245_plots_v2 <- change_plot(taxa = "bass", ssp="ssp245", habitat_only = TRUE)
bass_near_ssp245_v2 <- M_A_H(bass_near_ssp245_plots_v2$mainland, bass_near_ssp245_plots_v2$alaska, bass_near_ssp245_plots_v2$hawaii)

bass_near_ssp585_plots_v2 <- change_plot(taxa = "bass", ssp="ssp585", habitat_only = TRUE)
bass_near_ssp585_v2 <- M_A_H(bass_near_ssp585_plots_v2$mainland, bass_near_ssp585_plots_v2$alaska, bass_near_ssp585_plots_v2$hawaii)
 

# catfish
catfish_current_plots_v2 <- current_plot(taxa = "catfish", habitat_only = TRUE)
catfish_current_v2 <- M_A_H(catfish_current_plots_v2$mainland, catfish_current_plots_v2$alaska, catfish_current_plots_v2$hawaii)

catfish_near_ssp245_plots_v2 <- change_plot(taxa = "catfish", ssp="ssp245", habitat_only = TRUE)
catfish_near_ssp245_v2 <- M_A_H(catfish_near_ssp245_plots_v2$mainland, catfish_near_ssp245_plots_v2$alaska, catfish_near_ssp245_plots_v2$hawaii)

catfish_near_ssp585_plots_v2 <- change_plot(taxa = "catfish", ssp="ssp585", habitat_only = TRUE)
catfish_near_ssp585_v2 <- M_A_H(catfish_near_ssp585_plots_v2$mainland, catfish_near_ssp585_plots_v2$alaska, catfish_near_ssp585_plots_v2$hawaii)
 

# carp
carp_current_plots_v2 <- current_plot(taxa = "carp", habitat_only = TRUE)
carp_current_v2 <- M_A_H(carp_current_plots_v2$mainland, carp_current_plots_v2$alaska, carp_current_plots_v2$hawaii)

carp_near_ssp245_plots_v2 <- change_plot(taxa = "carp", ssp="ssp245", habitat_only = TRUE)
carp_near_ssp245_v2 <- M_A_H(carp_near_ssp245_plots_v2$mainland, carp_near_ssp245_plots_v2$alaska, carp_near_ssp245_plots_v2$hawaii)

carp_near_ssp585_plots_v2 <- change_plot(taxa = "carp", ssp="ssp585", habitat_only = TRUE)
carp_near_ssp585_v2 <- M_A_H(carp_near_ssp585_plots_v2$mainland, carp_near_ssp585_plots_v2$alaska, carp_near_ssp585_plots_v2$hawaii)
 

# tilapia
tilapia_current_plots_v2 <- current_plot(taxa = "tilapia", habitat_only = TRUE)
tilapia_current_v2 <- M_A_H(tilapia_current_plots_v2$mainland, tilapia_current_plots_v2$alaska, tilapia_current_plots_v2$hawaii)

tilapia_near_ssp245_plots_v2 <- change_plot(taxa = "tilapia", ssp="ssp245", habitat_only = TRUE)
tilapia_near_ssp245_v2 <- M_A_H(tilapia_near_ssp245_plots_v2$mainland, tilapia_near_ssp245_plots_v2$alaska, tilapia_near_ssp245_plots_v2$hawaii)

tilapia_near_ssp585_plots_v2 <- change_plot(taxa = "tilapia", ssp="ssp585", habitat_only = TRUE)
tilapia_near_ssp585_v2 <- M_A_H(tilapia_near_ssp585_plots_v2$mainland, tilapia_near_ssp585_plots_v2$alaska, tilapia_near_ssp585_plots_v2$hawaii)
 

# perch
perch_current_plots_v2 <- current_plot(taxa = "perch", habitat_only = TRUE)
perch_current_v2 <- M_A_H(perch_current_plots_v2$mainland, perch_current_plots_v2$alaska, perch_current_plots_v2$hawaii)

perch_near_ssp245_plots_v2 <- change_plot(taxa = "perch", ssp="ssp245", habitat_only = TRUE)
perch_near_ssp245_v2 <- M_A_H(perch_near_ssp245_plots_v2$mainland, perch_near_ssp245_plots_v2$alaska, perch_near_ssp245_plots_v2$hawaii)

perch_near_ssp585_plots_v2 <- change_plot(taxa = "perch", ssp="ssp585", habitat_only = TRUE)
perch_near_ssp585_v2 <- M_A_H(perch_near_ssp585_plots_v2$mainland, perch_near_ssp585_plots_v2$alaska, perch_near_ssp585_plots_v2$hawaii)
 

# trout
trout_current_plots_v2 <- current_plot(taxa = "trout", habitat_only = TRUE)
trout_current_v2 <- M_A_H(trout_current_plots_v2$mainland, trout_current_plots_v2$alaska, trout_current_plots_v2$hawaii)

trout_near_ssp245_plots_v2 <- change_plot(taxa = "trout", ssp="ssp245", habitat_only = TRUE)
trout_near_ssp245_v2 <- M_A_H(trout_near_ssp245_plots_v2$mainland, trout_near_ssp245_plots_v2$alaska, trout_near_ssp245_plots_v2$hawaii)

trout_near_ssp585_plots_v2 <- change_plot(taxa = "trout", ssp="ssp585", habitat_only = TRUE)
trout_near_ssp585_v2 <- M_A_H(trout_near_ssp585_plots_v2$mainland, trout_near_ssp585_plots_v2$alaska, trout_near_ssp585_plots_v2$hawaii)
 

# sturgeon
sturgeon_current_plots_v2 <- current_plot(taxa = "sturgeon", habitat_only = TRUE)
sturgeon_current_v2 <- M_A_H(sturgeon_current_plots_v2$mainland, sturgeon_current_plots_v2$alaska, sturgeon_current_plots_v2$hawaii)

sturgeon_near_ssp245_plots_v2 <- change_plot(taxa = "sturgeon", ssp="ssp245", habitat_only = TRUE)
sturgeon_near_ssp245_v2 <- M_A_H(sturgeon_near_ssp245_plots_v2$mainland, sturgeon_near_ssp245_plots_v2$alaska, sturgeon_near_ssp245_plots_v2$hawaii)

sturgeon_near_ssp585_plots_v2 <- change_plot(taxa = "sturgeon", ssp="ssp585", habitat_only = TRUE)
sturgeon_near_ssp585_v2 <- M_A_H(sturgeon_near_ssp585_plots_v2$mainland, sturgeon_near_ssp585_plots_v2$alaska, sturgeon_near_ssp585_plots_v2$hawaii)
 

```


Put it all together
```{r}

current <- ggarrange(trout_current, NULL, tilapia_current, NULL, catfish_current, widths = c(1, 0.04, 1, 0.04, 1), nrow=1, common.legend=TRUE)

ssp2_all <- ggarrange(trout_near_ssp245, NULL, tilapia_near_ssp245, NULL, 
                      catfish_near_ssp245, widths = c(1, 0.04, 1, 0.04, 1), nrow=1, common.legend=TRUE)

ssp2_farms <- ggarrange(trout_near_ssp245_v2, NULL, tilapia_near_ssp245_v2, NULL,
                      catfish_near_ssp245_v2, widths = c(1, 0.04, 1, 0.04, 1), nrow=1, common.legend=TRUE)

ssp5_all <- ggarrange(trout_near_ssp585, NULL, tilapia_near_ssp585, NULL, 
                      catfish_near_ssp585, widths = c(1, 0.04, 1, 0.04, 1), nrow=1, common.legend=TRUE)

ssp5_farms <- ggarrange(trout_near_ssp585_v2, NULL,  tilapia_near_ssp585_v2, NULL, 
                      catfish_near_ssp585_v2, widths = c(1, 0.04, 1, 0.04, 1), nrow=1, common.legend=TRUE)

ssp2 <- ggarrange(ssp2_farms, ssp2_all, nrow=2, common.legend = FALSE)
ssp5 <- ggarrange(ssp5_farms, ssp5_all, nrow=2, common.legend = FALSE)


pdf(file = here("figs/output/fig_3.pdf"), width = 8, height = 9)
ggarrange(current, NULL, ssp2, NULL, ssp5, nrow=5, heights = c(1, 0.1, 2, 0.06, 2), common.legend=FALSE)
dev.off()

myplot <- ggarrange(current, NULL, ssp2, NULL, ssp5, nrow=5, heights = c(1, 0.1, 2, 0.06, 2), common.legend=FALSE)
myplot
ggsave("figs/output/fig_3.png", width=8, height=9, units="in")

```


Put the remaining taxa together for supplement:
```{r}

current <- ggarrange(bass_current, NULL, carp_current, NULL, perch_current, NULL, sturgeon_current, widths = c(1, 0.04, 1, 0.04, 1, 0.04, 1), nrow=1, common.legend=TRUE)

ssp2_all <- ggarrange(bass_near_ssp245, NULL, carp_near_ssp245, NULL, 
                      perch_near_ssp245, NULL, sturgeon_near_ssp245, widths = c(1, 0.04, 1, 0.04, 1, 0.04, 1), nrow=1, common.legend=TRUE)

ssp2_farms <- ggarrange(bass_near_ssp245_v2, NULL, carp_near_ssp245_v2, NULL,
                     perch_near_ssp245_v2, NULL, sturgeon_near_ssp245_v2, widths = c(1, 0.04, 1, 0.04, 1, 0.04, 1), nrow=1, common.legend=TRUE)

ssp5_all <- ggarrange(bass_near_ssp585, NULL, carp_near_ssp585, NULL, 
                      perch_near_ssp585, NULL, sturgeon_near_ssp585, widths = c(1, 0.04, 1, 0.04, 1, 0.04, 1), nrow=1, common.legend=TRUE)

ssp5_farms <- ggarrange(bass_near_ssp585_v2, NULL,  carp_near_ssp585_v2, NULL, 
                      perch_near_ssp585_v2, NULL, sturgeon_near_ssp585_v2, widths = c(1, 0.04, 1, 0.04, 1, 0.04, 1), nrow=1, common.legend=TRUE)

ssp2 <- ggarrange(ssp2_farms, ssp2_all, nrow=2, common.legend = FALSE)
ssp5 <- ggarrange(ssp5_farms, ssp5_all, nrow=2, common.legend = FALSE)


# pdf(file = here("figs/output/fig_S1_new.pdf"), width = 8, height = 9)
# ggarrange(current, NULL, ssp2, NULL, ssp5, nrow=5, heights = c(1, 0.1, 2, 0.06, 2), common.legend=FALSE)
# dev.off()

myplot <- ggarrange(current, NULL, ssp2, NULL, ssp5, nrow=5, heights = c(1, 0.1, 2, 0.06, 2), common.legend=FALSE)
myplot
ggsave("figs/output/fig_S1.png", width=7, height=8, units="in")

```





OLD PLOTS BELOW, delete after we confirm we like the new idea!






function to make a 1 x 3 plot for each taxa of current and near-future projections for ssp2 and ssp5
```{r}

taxon_plot <- function(taxon){ #taxon = "carp"

## plot function start
taxa_plot_function <- function(rasterfile){ #rasterfile = current_raster_pt
  ggplot() +
 geom_tile(data = rasterfile, aes(x = x, y = y, fill = value)) +
 theme_classic() +
 scale_fill_gradientn(colours = current_cols, limit=c(0, 1)) +
 xlim(-4000000, 3000000) + ylim(100000, 6200000) +
 theme_void() + 
 theme(legend.position = "none",
       panel.border = element_rect(fill = NA, color = "lightgray", size = 2)) + # This line adds the border
 geom_sf(data=us_outline, fill=NA, color = "darkgray", size=2)
}
## plot function end  
  
# current plot
current <- sprintf("/home/shares/aquaculture/aquaculture_climatechange_usa/prep_data/freshwater/Tavg_taxa_prs/rescaled_yearly_crs/%s_Tavg_ssp245_2001-2020.tif", taxon)
current_raster <- terra::rast(current) 
current_raster <- terra::ifel(is.na(current_raster), 0, current_raster) * US
current_raster_pt <- as.data.frame(current_raster, xy=TRUE)
colnames(current_raster_pt) <- c("x", "y", "value")

current_plot <- taxa_plot_function(current_raster_pt)

# ssp2 plot
ssp2 <- sprintf("/home/shares/aquaculture/aquaculture_climatechange_usa/prep_data/freshwater/Tavg_taxa_prs/rescaled_yearly_crs/%s_Tavg_ssp245_2041-2060.tif", taxon)
ssp2_raster <- terra::rast(ssp2)
ssp2_raster <- terra::ifel(is.na(ssp2_raster), 0, ssp2_raster) * US
ssp2_raster_pt <- as.data.frame(ssp2_raster, xy=TRUE)
colnames(ssp2_raster_pt) <- c("x", "y", "value")

ssp2_plot <- taxa_plot_function(ssp2_raster_pt)

# ssp5 plot
ssp5 <- sprintf("/home/shares/aquaculture/aquaculture_climatechange_usa/prep_data/freshwater/Tavg_taxa_prs/rescaled_yearly_crs/%s_Tavg_ssp585_2041-2060.tif", taxon)
ssp5_raster <- terra::rast(ssp5)
ssp5_raster <- terra::ifel(is.na(ssp5_raster), 0, ssp5_raster) * US
ssp5_raster_pt <- as.data.frame(ssp5_raster, xy=TRUE)
colnames(ssp5_raster_pt) <- c("x", "y", "value")

ssp5_plot <- taxa_plot_function(ssp5_raster_pt)

ggarrange(current_plot, ssp2_plot, ssp5_plot, nrow=1)

}

```

Make plots for each taxa
```{r}
bass_plot <- taxon_plot("bass")
carp_plot <- taxon_plot("carp")
catfish_plot <- taxon_plot("catfish")
perch_plot <- taxon_plot("perch")
sturgeon_plot <- taxon_plot("sturgeon")
tilapia_plot <- taxon_plot("tilapia")
trout_plot <- taxon_plot("trout")

myplot <- ggarrange(NULL, bass_plot, NULL, catfish_plot, NULL, carp_plot, NULL, tilapia_plot, NULL, trout_plot, NULL, perch_plot, NULL, sturgeon_plot, nrow=14, common.legend = TRUE, heights=c(0.2, 1, 0.02, 1, 0.02, 1, 0.02, 1, 0.02, 1, 0.02, 1, 0.02, 1))
myplot
ggsave("figs/output/fig_S1.png", width=3.5, height=8, units="in")
