---
title: "Taxa biological pressures"
output: html_document
date: "2023-10-03"
editor_options: 
  chunk_output_type: console
---
This creates current maps for each taxa for sensitivity to thermal pressure model.
```{r}
library(terra)
library(tidyverse)
library(RColorBrewer)

current_cols <- c("#f0eada", brewer.pal("RdPu", n=9)[2:9])

US <- terra::rast(here("spatial/USA_raster.tif"))


```


function to make a 1 x 3 plot for each taxa of current model (1 month buffer) and more strict model (no buffer) and less strict model (2 month buffer)
```{r}

taxon_plot <- function(taxon){ #taxon = "carp"

## plot function start
taxa_plot_function <- function(rasterfile){ #rasterfile = current_raster_pt
  ggplot() +
 geom_tile(data = rasterfile, aes(x = x, y = y, fill = value)) +
 theme_classic() +
 scale_fill_gradientn(colours = current_cols, limit=c(0, 1)) +
 xlim(-4000000, 3000000) + ylim(100000, 6200000) +
 theme_void() + 
 theme(legend.position = "none",
       panel.border = element_rect(fill = NA, color = "lightgray", size = 2)) + # This line adds the border
 geom_sf(data=us_outline, fill=NA, color = "darkgray", size=2)
}
## plot function end  
  
# current plot
current <- sprintf("/home/shares/aquaculture/aquaculture_climatechange_usa/prep_data/freshwater/Tavg_taxa_prs/rescaled_yearly_crs/%s_Tavg_ssp245_2001-2020.tif", taxon)
current_raster <- terra::rast(current) 
current_raster <- terra::ifel(is.na(current_raster), 0, current_raster) * US
current_raster_pt <- as.data.frame(current_raster, xy=TRUE)
colnames(current_raster_pt) <- c("x", "y", "value")

current_plot <- taxa_plot_function(current_raster_pt)

# no_buffer
no_buffer <- sprintf("/home/shares/aquaculture/aquaculture_climatechange_usa/prep_data/freshwater/Tavg_taxa_prs/sensitivity/rescaled_yearly_crs/no_buffer_%s_Tavg_ssp245_2001-2020.tif", taxon)
no_buffer_raster <- terra::rast(no_buffer)
no_buffer_raster <- terra::ifel(is.na(no_buffer_raster), 0, no_buffer_raster) * US
no_buffer_raster_pt <- as.data.frame(no_buffer_raster, xy=TRUE)
colnames(no_buffer_raster_pt) <- c("x", "y", "value")

no_buffer_plot <- taxa_plot_function(no_buffer_raster_pt)

# extra buffer plot
extra_buffer <-  sprintf("/home/shares/aquaculture/aquaculture_climatechange_usa/prep_data/freshwater/Tavg_taxa_prs/sensitivity/rescaled_yearly_crs/two_month_buffer_%s_Tavg_ssp245_2001-2020.tif", taxon)
two_month_buffer_raster <- terra::rast(extra_buffer)
two_month_buffer_raster <- terra::ifel(is.na(two_month_buffer_raster), 0, two_month_buffer_raster) * US
two_month_buffer_raster_pt <- as.data.frame(two_month_buffer_raster, xy=TRUE)
colnames(two_month_buffer_raster_pt) <- c("x", "y", "value")

two_month_buffer_plot <- taxa_plot_function(two_month_buffer_raster_pt)

ggarrange(current_plot, no_buffer_plot, two_month_buffer_plot, nrow=1)

}

```

Make plots for each taxa
```{r}
bass_plot <- taxon_plot("bass")
carp_plot <- taxon_plot("carp")
catfish_plot <- taxon_plot("catfish")
perch_plot <- taxon_plot("perch")
sturgeon_plot <- taxon_plot("sturgeon")
tilapia_plot <- taxon_plot("tilapia")
trout_plot <- taxon_plot("trout")

myplot <- ggarrange(NULL, bass_plot, NULL, catfish_plot, NULL, carp_plot, NULL, tilapia_plot, NULL, trout_plot, NULL, perch_plot, NULL, sturgeon_plot, nrow=14, common.legend = TRUE, heights=c(0.2, 1, 0.02, 1, 0.02, 1, 0.02, 1, 0.02, 1, 0.02, 1, 0.02, 1))
myplot
ggsave("figs/output/fig_S2.png", width=3.5, height=8, units="in")
