---
title: "Likely aquaculture locations for each taxa"
output: html_document
date: "2023-09-11"
editor_options: 
  chunk_output_type: console
---

Goal: determine the most probable location of farm locations for each taxa based on state level data for each species, county level data indicating presence/absence of aquaculture (not species specific), and our maps of pressures for each taxa.
```{r setup, include=FALSE}
library(here)
library(terra)
library(tidyverse)
library(sf)
library(raster)

#raster template
template <- raster()
template <- raster(xmn=-6500000, xmx=2500000, ymn=100000, ymx=6200000)

crs(template) <- CRS('+init=EPSG:5070')
res(template) <- 5000

#writeRaster(template, here("spatial/template.tif"))
```

Get the mapping data!
Created in "aquaculture_location/location_of_aquaculture.Rmd"
```{r}
map_counties <- st_read("spatial/state_county_shp/state_county_shp.gpkg") %>%
  mutate(STATEFP = as.numeric(STATEFP))
fips <- read_csv(here("spatial/county_state_fips_ID.csv")) %>%
  mutate(State = toupper(state_name)) %>%
  dplyr::select(state_fips, State) %>%
  unique()

```

Make a raster for each fish taxa with the states that it is known to be reared in. Halley provided the file titled: max_value_census.csv.  
```{r}

# prepare the data

location <- read.csv(here("aquaculture_location/data/max_value_census.csv")) 
names(location) <- tolower(names(location))
names(location)[1] <- "State"


location <- left_join(fips, location, by = "State")

location <- location %>%
  rename(STATEFP=state_fips) %>%
  mutate(across(bass:trout, ~ case_when(
    . >= 0 ~ 1L,
    is.na(.) ~ 0L,
    TRUE ~ NA_integer_
  )))

setdiff(map_counties$STATEFP, location$STATEFP)
setdiff(location$STATEFP, map_counties$STATEFP)

map_states <- left_join(map_counties, location, by="STATEFP") %>%
  mutate(land = 1)
land_raster <- fasterize::fasterize(map_states, template, field="land")
#plot(land_raster)
```

Loop through each taxa.
```{r}
taxas = c("carp", "catfish", "perch", "sturgeon", "tilapia", "trout", "bass")

for(taxa in taxas){ # taxa = taxas[2]
  
    raster_states <- fasterize::fasterize(map_states, template, field=taxa)
    raster_states[is.na(raster_states)] <- 0
    raster_states <- raster_states * land_raster
    writeRaster(raster_states, sprintf("/home/shares/aquaculture/aquaculture_climatechange_usa/prep_data/freshwater/mapping_aquaculture/taxa_state_maps/%s.tif", taxa), overwrite=TRUE)
}

```

Make maps that multiply the state/taxa maps by the current location of aquaculture maps (not taxa specific) and the cumulative pressure rasters for each taxa.

```{r}

# raster of state IDs
state_id <- fasterize::fasterize(map_states, template, field="STATEFP")
state_id <- rast(state_id)

# get the general map describing, by county, the location of aquaculture
county_aqua <- rast("/home/shares/aquaculture/aquaculture_climatechange_usa/prep_data/freshwater/mapping_aquaculture/aquaculture_location.tif")
county_aqua <- ifel(county_aqua >= 1, 1, 0)

# loop through all the taxa:

taxas = c("carp", "catfish", "perch", "sturgeon", "tilapia", "trout", "bass")

for (taxa in taxas){ # taxa = taxas[1]
  state_taxa <- rast(sprintf("/home/shares/aquaculture/aquaculture_climatechange_usa/prep_data/freshwater/mapping_aquaculture/taxa_state_maps/%s.tif", taxa)) 
  
  pressure <- rast(sprintf("/home/shares/aquaculture/aquaculture_climatechange_usa/prep_data/freshwater/combined_pressures/taxa_pressures/current_ssp245_%s.tif", taxa))
  
# pressure2 <- rast(sprintf("/home/shares/aquaculture/aquaculture_climatechange_usa/prep_data/freshwater/combined_pressures/taxa_pressures/current_ssp585_%s.tif", taxa))
  
  pressure <- ifel(pressure <= 0.75, 1, 0)
#  pressure2 <- ifel(pressure2 <= 0.75, 1, 0)
#  plot(pressure - pressure2)
    likely <- state_taxa * pressure * county_aqua
  
  modeled_obs <- zonal(likely, state_id, fun=sum, na.rm=TRUE) 
  
  modeled_obs <- as.data.frame(modeled_obs) %>%
    rename(STATEFP = layer) %>%
    mutate(value = ifelse(.[[taxa]]>0, 1, 0)) %>%
    dplyr::select(STATEFP, value)

 observed <- location  %>%
    dplyr::select(STATEFP, State, any_of(taxa))

 compare <- left_join(observed, modeled_obs, by="STATEFP") %>%  
 filter(.data[[taxa]] == 1 & value == 0) %>%
   dplyr::select(State) %>%
   pull()
 
 cat("For ", taxa, " = these states were missing: ", compare, "\n")

 writeRaster(likely, sprintf("/home/shares/aquaculture/aquaculture_climatechange_usa/prep_data/freshwater/mapping_aquaculture/estimated_farm_locations/%s.tif", taxa), overwrite=TRUE)
 
}

```

Make figures of where each taxa has the highest probability of being farmed.
```{r}
for(taxa in taxas){
  png(sprintf(here("aquaculture_location/output/likely_location_%s.png"), taxa), width=800, height=600)
plot(raster::raster(sprintf("/home/shares/aquaculture/aquaculture_climatechange_usa/prep_data/freshwater/mapping_aquaculture/estimated_farm_locations/%s.tif", taxa)), main=taxa, legend=FALSE)
  dev.off()
}
```

