---
title: "location_of_aquaculture"
author: "Cat Fong"
date: '2022-07-06'
editor_options: 
  chunk_output_type: console
---

```{r}
library(sf)
library(sp)
library(dplyr)
library(tidyverse)
library(stringr)
library(ggplot2)
library(fasterize)
library(ggmap)
library(raster)
library(rgdal)
library(climateStability)
library(viridis)
library(RColorBrewer)
library(here)

```
##Methods for each piece

#Get the map data

1) Import all data on location
-Establishments (USBLS Census)
-Water use (USGS) (aquaculture is rows 224- 235 in the water file)
-Farms (USDA Census)

## Establishments (USBLS Census)
```{r Location of Aquaculture 1}

establishments_raw <- read.csv("/home/shares/aquaculture/aquaculture_climatechange_usa/raw_data/location_of_aquaculture/establishments.csv")

keep <- c("fips", "annual_avg_estabs_count")

establishments <- establishments_raw[keep]

```

## Water (USGS)

```{r}
water <- read.csv("/home/shares/aquaculture/aquaculture_climatechange_usa/raw_data/location_of_aquaculture/water.csv", stringsAsFactors = FALSE)

water_2015 <- subset(water, year=="2015")

water_cols <- grep("fips|Aquaculture.total.self.supplied.withdrawals..fresh..in.Mgal.d", names(water_2015))
water_aq <- water_2015[water_cols]


#write.table(water_aq, "water_aq.csv", sep=",")
#water_aq<-read.csv("/home/shares/aquaculture/aquaculture_climatechange_usa/raw_data/location_of_aquaculture/water_aq.csv", stringsAsFactors = FALSE)

water_aq$fips<-str_pad(water_aq$fips,5, pad = "0")

```

## Farms (USDA Census)

```{r}
farms <- read.csv("/home/shares/aquaculture/aquaculture_climatechange_usa/raw_data/location_of_aquaculture/Farms.csv")

farms <- subset(farms, Domain=="TOTAL")

fips_and_zips <- read.csv("/home/shares/aquaculture/aquaculture_climatechange_usa/raw_data/location_of_aquaculture/fips_and_zips.csv") %>% unique()

farms <- left_join(farms, fips_and_zips, by="Zip.Code")

farms$fips <- str_pad(farms$fips,5, pad = "0")
farms <- farms %>%
  dplyr::group_by(fips) %>%
  dplyr::summarize(Value = sum(Value, na.rm=TRUE))

```

2) Import empty maps and clean data
-Zip codes: https://www.census.gov/programs-surveys/geography/guidance/geo-areas/zctas.html
-Fips codes: https://www.census.gov/programs-surveys/geography/guidance/geo-identifiers.html

empty_zips contains the zip code shape files
empty_counties contains the county level shape files. I elected to use the 500k because this is the finest resolution (1:500,000). 
In the data file, STATEFP is the 2 digit county code for the state or territory while COUNTYFP is the 3 digit county code within the state. Concatanation would result in the FIPS code. 

Clean up empty map data; Filter by the STATEFP associated with the states(ie remove territories)
  American Samoa,	60
  Guam,	66
  Northern Mariana Islands,	69
  Puerto Rico,	72
  Virgin Islands,	78
  district of columbia, 11

Concatanate

in 5070 crs

```{r Location of Aquaculture 2}

#empty_zips <- st_read("/home/shares/aquaculture/aquaculture_climatechange_usa/empty_maps/cb_2018_us_zcta210_500k/cb_2018_us_zcta510_500k.shp")

empty_counties <- st_read("/home/shares/aquaculture/aquaculture_climatechange_usa/empty_maps/cb_2018_us_county_500k/cb_2018_us_county_500k.shp")

empty_counties <- empty_counties %>%
filter(!(STATEFP %in% c("60", "66", "69", "72", "78", "11" ) )) %>%
  mutate(ALAND = round(ALAND,0),
         AWATER = round(AWATER,0))

fips <- paste0(empty_counties$STATEFP, empty_counties$COUNTYFP)

empty_counties <- cbind(fips, empty_counties)

empty_counties <- empty_counties %>%
  st_transform(crs = 5070)
st_write(empty_counties, "spatial/state_county_shp/state_county_shp.gpkg")
```

3)Merge the county level data with the estimates of where aquaculture
```{r Location of Aquaculture 3}

## checking data to make sure merges go well
est_extras <- setdiff(establishments$fips, empty_counties$fips)
tmp <- filter(establishments_raw, fips %in% est_extras) # states, other(towns/cities), puerto rico

wat_extras <- setdiff(water_aq$fips, empty_counties$fips)
filter(water, fips %in% wat_extras) # cutting: 46113   SD Shannon County
filter(water, fips %in% "2270") # 2270 AK Wade Hampton Census Area

farms_extras <- setdiff(farms$fips, empty_counties$fips)


# joining 3 datasources to get most complete data.
aq <- full_join(establishments, water_aq, by="fips")
aq <- full_join(aq, farms, by="fips")


all_data<-left_join(empty_counties, aq, by="fips")


```
4) make aquaculture data binary

```{r Location of Aquaculture 4}
all_data$annual_avg_estabs_count[all_data$annual_avg_estabs_count >= 1] <- 1
all_data$annual_avg_estabs_count[all_data$annual_avg_estabs_count < 1] <- 0
all_data$annual_avg_estabs_count[is.na(all_data$annual_avg_estabs_count)] <- 0 #are NA's zero?

all_data$Value[all_data$Value >= 1] <- 1
all_data$Value[all_data$Value < 1] <- 0
all_data$Value[is.na(all_data$Value)] <- 0 #are NA's zero?

all_data$Aquaculture.total.self.supplied.withdrawals..fresh..in.Mgal.d[all_data$Aquaculture.total.self.supplied.withdrawals..fresh..in.Mgal.d >= 1] <- 1

all_data$Aquaculture.total.self.supplied.withdrawals..fresh..in.Mgal.d[all_data$Aquaculture.total.self.supplied.withdrawals..fresh..in.Mgal.d < 1] <- 0

all_data$Aquaculture.total.self.supplied.withdrawals..fresh..in.Mgal.d[is.na(all_data$Aquaculture.total.self.supplied.withdrawals..fresh..in.Mgal.d)] = 0 #are NA's zero?

all_data <- all_data %>%
  mutate(fips = as.numeric(fips))

#saveRDS(all_data, "/home/shares/aquaculture/aquaculture_climatechange_usa/prep_data/freshwater/mapping_aquaculture/all_data.rds")

```

5) Create empty raster and to the fasterize thing
```{r Location of Aquaculture 5}

x <- raster()
x <- raster(xmn=-6500000, xmx=2500000, ymn=100000, ymx=6200000)

crs(x) <- crs(all_data)
res(x) <- 5000


fips_raster<-fasterize(all_data, x, field="fips")
#plot(fips_raster)

```

6) Make some maps

```{r Location of Aquaculture 6}

establishment_raster <- fasterize(all_data, x, field="annual_avg_estabs_count")
plot(establishment_raster, box = FALSE, axes = FALSE,main="Establishments")

farms_raster <- fasterize(all_data, x, field="Value")
plot(farms_raster,box = FALSE, axes = FALSE, main="Farms")

water_raster <- fasterize(all_data, x, field="Aquaculture.total.self.supplied.withdrawals..fresh..in.Mgal.d")
plot(water_raster, box = FALSE, axes = FALSE, main="Water")

all <- establishment_raster + farms_raster + water_raster

plot(all,
     box = FALSE,
     axes = FALSE,
     main = "Aquaculture")

writeRaster(all, "/home/shares/aquaculture/aquaculture_climatechange_usa/prep_data/freshwater/mapping_aquaculture/aquaculture_location.tif", overwrite=TRUE)

tmp <- raster("/home/shares/aquaculture/aquaculture_climatechange_usa/prep_data/freshwater/mapping_aquaculture/aquaculture_location.tif")
```