---
title: "zonal_stats"
output: html_document
date: "2022-07-29"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}

library(raster)
library(fasterize)
library(sf)
library(tidyverse)
library(here)

```

## Clean fips ID data
Source: U.S. Census Bureau, Population Division	
Internet Release Date: May 2018	
```{r}
fips <- read_csv(here("spatial/all-geocodes-v2017.csv"))
table(fips$summary_level)

fips %>%
  filter(summary_level == "061") #towns

fips %>%
  filter(summary_level == "162") #cities

fips %>%
  filter(summary_level == "170") #????


states <- fips %>%
  filter(summary_level=="040") %>%
  select(state_fips, state = name)

county <- fips %>%
  filter(summary_level == "050") %>%
  select(state_fips, county_fips, county = name)

region_id <- left_join(county, states, by="state_fips") %>%
  mutate(fips = paste0(state_fips, county_fips))
```

##County polygon files

Import empty maps and clean data
-Fips codes: https://www.census.gov/programs-surveys/geography/guidance/geo-identifiers.html

empty_counties contains the county level shape files. I elected to use the 500k because this is the finest resolution (1:500,000). 
In the data file, STATEFP is the 2 digit county code for the state or territory while COUNTYFP is the 3 digit county code within the state. Concatanation would result in the FIPS code. 

Clean up empty map data; Filter by the STATEFP associated with the states(ie remove territories)
  American Samoa,	60
  Guam,	66
  Northern Mariana Islands,	69
  Puerto Rico,	72
  Virgin Islands,	78
  district of columbia, 11

Concatanate

in 5070 crs

```{r Location of Aquaculture 2}

map_counties <- st_read("/home/shares/aquaculture/aquaculture_climatechange_usa/empty_maps/cb_2018_us_county_500k/cb_2018_us_county_500k.shp")

map_counties <- map_counties[!map_counties$STATEFP  %in% c("60", "66", "69", "72", "78", "11" ),]

map_counties$fips <- paste0(map_counties$STATEFP, map_counties$COUNTYFP)

map_counties <- map_counties %>%
  st_transform(crs = 5070)

```

## Join data
Need to get state names!

```{r}
# check they all match up

setdiff(map_counties$fips, region_id$fips)
missing_map <- setdiff(region_id$fips, map_counties$fips)

tmp <- filter(region_id, fips %in% missing_map)
unique(tmp$state) #District of Colombia and Puerto Rico, this is fine

map_counties_id <- map_counties %>%
  left_join(region_id, by = "fips") %>%
  mutate(state_fips = as.numeric(state_fips)) %>%
  mutate(fips = as.numeric(fips)) %>%
  mutate(us_id = 1)

county_id_csv <- as_tibble(map_counties_id) %>%
  select(state_fips, county_fips=fips, county_name=NAME, state_name=state, area_land=ALAND, area_water=AWATER)
write_csv(county_id_csv, here("spatial/county_state_fips_ID.csv"))

```


Make some county and state and US rasters.
```{r}
template <- raster("/home/shares/aquaculture/aquaculture_climatechange_usa/prep_data/freshwater/mapping_aquaculture/aquaculture_location.tif")

state_ID <- fasterize(map_counties_id, template, field="state_fips")
county_ID <- fasterize(map_counties_id, template, field="fips")
usa_ID <- fasterize(map_counties_id, template, field="us_id")
writeRaster(state_ID, here("spatial/state_fips_raster.tif"), overwrite=TRUE)
writeRaster(county_ID, here("spatial/county_fips_raster.tif"), overwrite=TRUE)
writeRaster(usa_ID, here("spatial/USA_raster.tif"), overwrite=TRUE)
```